---
title: "ecoli_analysis"
output: html_document
---

# Heatmap and Correlation plot

Import required packages
```{r}
library(readxl)
library(tidyverse)
library(corrplot) # to create correlation matrix plot
library(heatmap3)
library(pheatmap)
```

## Heatmap and Hierarchical Figure 5

Import Data
```{r}
data_heat <- read_excel("data_E.coli - Coded2(V2).xlsx",sheet = 1)
```

Label Sample Source column
```{r}
data_heat$`Sample Source` <- as.factor (ifelse(data_heat$`Sample Source`==0,"Environmental","Clinical"))
```

Separate row annotation
```{r}
ann_row <- data.frame(data_heat[,c("Sampling_point", "Sample Source", "Hospital")])
```

Include sample Ids as row name
```{r}
rownames(ann_row) <- data_heat$Isolate
```

Convert to factor
```{r}
ann_row[,1]<-as.factor(ann_row[,1])
ann_row[,2]<-as.factor(ann_row[,2])
```

Change the Column name
```{r}
colnames(ann_row)<-c("Sampling Point", "Sample Source", "Hospital")
```

```{r}
rownames(ann_row) <- sub("^EC-", "LEH-", rownames(ann_row))
```

Select data
```{r}
mat_work <- as.matrix(data_heat[,c(7:25)])
rownames(mat_work) <- data_heat$Isolate
rownames(mat_work) <- sub("^EC-", "LEH-", rownames(mat_work))
```


```{r}
ann_color<-list("Sample Source"=c(Clinical='#464DE5',Environmental="yellow"))
```


```{r}
# Define the output file and set the dimensions
#png("./Outputs/heatmap.png", height = 1400, width = 500)

# Generate the heatmap
pheatmap(
  mat = mat_work,
  scale = "none",
  cluster_cols = FALSE,
  clustering_distance_rows = "binary",
  clustering_method = "ward.D2",
  col = colorRampPalette(c("Green", "Red"))(1024),
  annotation_row = ann_row,
  annotation_colors = ann_color,
  legend = FALSE,
  cellwidth = 11,
  cellheight = 11,
  margins = c(40, 20)
)

# Close the device
#dev.off()
```




## Correlation matrix of phenotype Figure 6




Import data
```{r}
data_cor <- read_excel("data_E.coli.xlsx", sheet = "Sheet2")
```

Select required column for correlation matrix
```{r}
data_cor <- data_cor[,c(7:32,33)]
```

Create a correlation matrix
The cor() function will compute pairwise correlations between the selected columns of the data frame.
```{r}
corr <- cor(data_cor)
```

Create correlation matrix plot
```{r}
corrplot(corr,type = "upper", diag = F, tl.col="black",insig = "blank")
```


